A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information

### 2017-08-19 analysis/paper/paper.Rmd
update plots

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

date_plot <- 
  dates %>% 
  group_by(yyyy_date) %>% 
  tally() %>% 
  ggplot(aes( yyyy_date, n)) +
  geom_col(fill = datacentre) +
  scale_y_log10() +
  theme_new() +
  xlim(c(1920,2017)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite")

date_plot <- 
  dates %>% 
  group_by(yyyy_date) %>% 
  tally() %>% 
  ggplot(aes( yyyy_date, n)) +
  geom_col() +
  scale_y_log10() +
  theme_new() +
  xlim(c(1920,2017)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite")

years_of_interest <- c(1957, 1990, 2009)
ymax <- 15000
date_plot + 
  geom_vline(xintercept = years_of_interest, colour = "red") +
  annotate("text", 
           x = years_of_interest + 4,
           y = ymax,
           label = years_of_interest) +
  scale_y_log10(limits = c(1, ymax))

# library(plotly)
# ggplotly(date_plot)

@@

 years_of_interest <- c(1990, 2009)
 ymax <- 10000
date_plot2 <- 
  dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre_symbol
             )) +
  geom_bar() +
  guides(fill=FALSE) +
  scale_fill_viridis(discrete = TRUE) +
  theme_new() +
  xlim(c(1950,2017)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = years_of_interest, colour = "red") +
  annotate("text", 
           x = years_of_interest + 2,
           y = ymax,
           label = years_of_interest)

@@

Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data). The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY). Although the earliest data dates to `r first_date`, most of the data was collected after 1957. The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable. There is a substantial increase in 2009 when the DANS-EASY

@@

@@

# why so many in 2010? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch

dates %>% 
  group_by(yyyy_date, datacentre) %>% 
   tally(sort = TRUE)  %>% 
   top_n(n=1) %>% 
  View

@@

why_2008 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2008)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)
```

```{r eval=TRUE}
# panel of plots
# combine plots for lang and data center
datacentre_and_langugage_plot <- 
  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
                                      xmin = 1.15, 
                                      xmax = 6.15,
                                      ymin = 10000,
                                      ymax = 40000
  )

@@

>>>>

---
title: "A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information"
author:
  - "Ben Marwick"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).

The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation

@@

@@

 years_of_interest <- c(2009)
 ymax <- 10000

date_plot <- 
  dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_viridis(discrete = TRUE,
                     breaks = top_data_sources[1:4],
                     labels = top_data_sources[1:4]) +
  theme_new() +
  theme(legend.position = c(0.3, 0.75)) +
  xlim(c(1950,2017)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = years_of_interest, colour = "red") +
  annotate("text", 
           x = years_of_interest + 2,
           y = ymax,
           label = years_of_interest)

@@

Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data). The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY). Although the earliest data dates to `r first_date`, most of the data was collected after 1957. The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize. There is a substantial increase in 2009 when the DANS-EASY

@@

# show plots
datacentre_plot
allocator_plot
creator_plot
language_plot
contributor_plot
rights_plot
publisher_plot

@@

# why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch

top_per_year <- 
dates %>% 
  group_by(yyyy_date, datacentre) %>% 
   tally(sort = TRUE)  %>% 
   top_n(n=1)

@@

why_2008 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2008)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

# in 2009, among the Dutch, what is going on?
dutch_2009 <- 
dc_arch %>% 
  filter(yyyy_date == 2009,
         datacentre == "DELFT.EASY - DANS-EASY") %>% 
  group_by(publisher) %>% 
  tally(sort=TRUE)

# how about in ADS?
ads_2009 <- 
dc_arch %>% 
  filter(yyyy_date == 2009,
         grepl("ADS", datacentre)) %>% 
  group_by(creator) %>% 
  tally(sort=TRUE)

# compare to 2009
dutch_2008 <- 
dc_arch %>% 
  filter(yyyy_date == 2008,
         datacentre == "DELFT.EASY - DANS-EASY") %>% 
  group_by(publisher) %>% 
  tally(sort=TRUE) 
```

```{r}
# panel of plots
# combine plots for lang and data center
datacentre_and_langugage_plot <- 
  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
                                      xmin = 1.15, 
                                      xmax = 6.15,
                                      ymin = 10000,
                                      ymax = 40000
  )

@@

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE, eval = TRUE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r eval = TRUE}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```
